---
title: 'Lab 1:'
output:
  html_document:
    df_print: paged
---

# Introduction

This lab focuses on text processing in R.

First we load the `datamining` package and the data for the lab.

```{r}
library(datamining)
load("lab1_docs.rda")
```

# Processing a text file

Using the functions described below, read the file `politics3.txt` into R. What is the 57th word
in the document?

# The full dataset

1 The full dataset is in a table called `doc`, containing 9 posts to `talk.politics.misc` and 9 posts
to `talk.religion.misc`. The `dim` function will tell you the size of the table:

```{r}
dim(doc)
```

A vector `doc.labels` is also defined.

2. According to `doc`, how many times does ‘the’ occur in politics3! Give a command to select
the answer from doc. (Note that both the rows and columns have names.)

# Document similarity
3. In preparation for computing distances, remove the words which occur only once, and weight
the remaining words by inverse document frequency (IDF). These two commands do it:
```{r eval=FALSE}
doc = remove.singletons(doc)
doc = idf.weight(doc)
```
`doc` is modified in place.
4. Compute a distance matrix between all documents, using IDF and without normalizing for
document length. Save the first five rows in your Word file, for use in the homework. You can
also write the entire table to a file, via
```{r eval=FALSE}
write .array(d,file="d.txt")
```
5. Compute a distance matrix where the weighted word counts have been normalized by the doc-
ument total. Save the first five rows.
6. Compute a distance matrix where the weighted word counts have been normalized by the doc-
ument’s Euclidean length. Save the first five rows.
7. Pick one of the above matrices and make a multidimensional scaling visualization of it. Save the plot.

# Reading a document into R

If the Desktop has a file called “politics3.txt”, you can read it into R via
```{r}
txt <- read_doc("politics3.txt")
```

This makes `txt` into a vector of words. The nth word can obtained by typing `txt[n]`. 
The function `read.doc` removes the message header, removes all punctuation and capitalization, and converts all numbers to the hash symbol #. Typing `table(txt)` gives the bag of words representation.

Selecting from a table or matrix When a table has two dimensions, like doc does, you select from
it by giving a row name or number and a column name or number, e.g. `doc[2,4]` or `doc[4,"the"]`. An
entire row or column can be selected by leaving out the index, e.g. `doc["politics3",]` or `doc[, 4]`.

# Computing distances 

If `x` has your (normalized and weighted) word counts, type
```{r eval=FALSE}
d <- distances(x)
```
and `d` will become a matrix of distances.

# Normalizing documents 

To divide each document’s count vector by its sum:
```{r}
x <- div_by_sum(doc)
```
The result is put into `x` so that the original doc can still be used. To instead divide each document's count vector by its Euclidean length:
```{r}
x <- div_by_euc_length(doc)
```

# Multidimensional scaling
If $d$ is a distance matrix and `doc.labels` contains the document labels
(for color-coding) then we can use the following function:
```{r eval=FALSE}
mds(d, doc.labels)
```
This will open a window for the plot. Using the menu, select Windows -> Tile to arrange the sub-
windows. You can also do multidimensional scaling with three dimensions instead of two, so that
distances can be represented more accurately.
mds(d,doc.labels,k=3)
This gives a 3D scene that you can rotate with the mouse.
