---
title: 'Lab 9: Linear Regression'
output:
  html_document:
    df_print: paged
---

# Introduction

This is Lab 9 on Data Mining. It is a simple introduction to the practicalities of linear regression.

First we load the `datamining` package and the data for the lab.
```{r warning=FALSE}
devtools::load_all()
load('lab9.rda')
```

# The dataset

The dataset is 196 weeks of grocery sales for a store. The variables are:

| Variable | Description |
| --- | --- |
| Price.1 | DOLE PINEAPPLE ORANG 64 OZ |
| Price.2 | FIVE ALIVE CTRUS BEV 64 OZ
| Price.3 | HH FRUIT PUNCH 64 OZ
| Price.4 | HH ORANGE JUICE 64 OZ
| Price.5 | MIN MAID O J CALCIUM 64 OZ
| Price.6 | MIN MAID O J PLASTIC 96 OZ
| Price.7 |  MM PULP FREE OJ 64 OZ
| Price.8 | SUNNY DELIGHT FLA CI 64 OZ
| Price.9 | TREE FRESH O J REG 64 OZ
| Price.10 | TROP PURE PRM HOMEST 64 OZ
| Price.l1 | TROP SB HOMESTYLE OJ 64 OZ
| Sold.4 | Number of units sold for HH ORANGE JUICE 64 OZ

It is stored in a matrix $x$. Look at the first few rows via x[1:3,].

```{r view_x}
head(x)
```


# Standardizing

First we make a copy of `x` and put the response firt in the `data.frame`. Then
we transform `Sold.4` appropriately and standardize all variables to have zero 
mean and unit variance.

```{r standardizing}
xx <- x[, c(ncol(x), 1:ncol(x)-1)]
hist(xx$Sold.4)
xx$Sold.4 <- log(xx$Sold.4)
xx <- as.data.frame(scale(xx))
hist(xx$Sold.4)
```

# Adding predictors

We construct a linear model to predict `Sold.4` as a function of Price.


```{r}
fit <- lm(Sold.4 ~ ., data=xx)
summary(fit)
```

Then we plot all variables versus the residuals of this model.

```{r all_residuals}
predict_plot(fit, xx)
```

The important predictors appear to be `Price.4`, `Price.5`, `Price.7` and `Price.11`.

We pick one of the important predictors from the last step and include it in a
new model.

```{r fit_price_4}
fit_4 <- lm(Sold.4 ~ Price.4, data=xx)
summary(fit_4)
```

```{r fit_4_residuals}
predict_plot(fit_4, xx)
```

The same predictors as before seem important with `Price.11` the most important.

Next we add `Price.11` to the model.

```{r fit_price_4_11}
fit_4_11 <- lm(Sold.4 ~ Price.4 + Price.11, data=xx)
summary(fit_4_11)
```

```{r fit_4_11_residuals}
predict_plot(fit_4_11, xx)
```

We keep adding predictors until no more predictors seem useful.

```{r fit_price_4_11_5_7}
fit_4_11_5_7 <- lm(Sold.4 ~ Price.4 + Price.11 + Price.5 + Price.7, data=xx)
summary(fit_4_11_5_7)
```

# Automatic selection

Starting again from a model with only `Price.4`, we use step to add predictors.

```{r step}
fit_step <- step(fit_4, formula(xx))
```

The automatically generated model is the same as that generated manually.

Finally we make a partial residual plot for the model from `step`.

```{r partial_plots}
predict_plot(fit_step, partial=TRUE)
```

Residual and partial residual plots are also available in the function 
`termplot` in the `stats` package.

```{r termplot}
library(stats)
par(mfrow = c(2, 2))
termplot(fit_step, partial.resid = TRUE, ask = FALSE)
par(mfrow = c(1, 1))
```

Note that the partial residuals for `Sold.4` are shown using `predict_plot` 
whereas `termplot` uses the partial residuals for each predictor on the vertical
axis.
